---
title: "Mini Data-Analysis Deliverable 1"
output:
  github_document
---

# Welcome to your (maybe) first-ever data analysis project!

And hopefully the first of many. Let's get started:

1.  Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package by typing the following into your **R terminal**:

<!-- -->

    install.packages("devtools")
    devtools::install_github("UBC-MDS/datateachr")

2.  Load the packages below.

```{r}
library(datateachr)
library(tidyverse)
```

3.  Make a repository in the <https://github.com/stat545ubc-2022> Organization. You will be working with this repository for the entire data analysis project. You can either make it public, or make it private and add the TA's and Lucy as collaborators. A link to help you create a private repository is available on the #collaborative-project Slack channel. 

# Instructions

## For Both Milestones

-   Each milestone is worth 45 points. The number of points allocated to each task will be annotated within each deliverable. Tasks that are more challenging will often be allocated more points.

-   10 points will be allocated to the reproducibility, cleanliness, and coherence of the overall analysis. While the two milestones will be submitted as independent deliverables, the analysis itself is a continuum - think of it as two chapters to a story. Each chapter, or in this case, portion of your analysis, should be easily followed through by someone unfamiliar with the content. [Here](https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/) is a good resource for what constitutes "good code". Learning good coding practices early in your career will save you hassle later on!

## For Milestone 1

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-1.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work below --->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to the mini-analysis GitHub repository you made earlier, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 45 points: 43 for your analysis, 1 point for having your Milestone 1 document knit error-free, and 1 point for tagging your release on Github.

# Learning Objectives

By the end of this milestone, you should:

-   Become familiar with your dataset of choosing
-   Select 4 questions that you would like to answer with your data
-   Generate a reproducible and clear report using R Markdown
-   Become familiar with manipulating and summarizing your data in tibbles using `dplyr`, with a research question in mind.

# Task 1: Choose your favorite dataset (10 points)

The `datateachr` package by Hayley Boyce and Jordan Bourak currently composed of 7 semi-tidy datasets for educational purposes. Here is a brief description of each dataset:

-   *apt_buildings*: Acquired courtesy of The City of Toronto's Open Data Portal. It currently has 3455 rows and 37 columns.

-   *building_permits*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 20680 rows and 14 columns.

-   *cancer_sample*: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.

-   *flow_sample*: Acquired courtesy of The Government of Canada's Historical Hydrometric Database. It currently has 218 rows and 7 columns.

-   *parking_meters*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 10032 rows and 22 columns.

-   *steam_games*: Acquired courtesy of Kaggle. It currently has 40833 rows and 21 columns.

-   *vancouver_trees*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 146611 rows and 20 columns.

**Things to keep in mind**

-   We hope that this project will serve as practice for carrying our your own *independent* data analysis. Remember to comment your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. As you advance in the project, prompts and hints to do this will be diminished - it'll be up to you!

-   Before choosing a dataset, you should always keep in mind **your goal**, or in other ways, *what you wish to achieve with this data*. This mini data-analysis project focuses on *data wrangling*, *tidying*, and *visualization*. In short, it's a way for you to get your feet wet with exploring data on your own.

And that is exactly the first thing that you will do!

1.1 Out of the 7 datasets available in the `datateachr` package, choose **4** that appeal to you based on their description. Write your choices below:

**Note**: We encourage you to use the ones in the `datateachr` package, but if you have a dataset that you'd really like to use, you can include it here. But, please check with a member of the teaching team to see whether the dataset is of appropriate complexity. Also, include a **brief** description of the dataset here to help the teaching team understand your data.

<!-------------------------- Start your work below ---------------------------->

1: apt_buildings

2: cancer_sample

3: steam_games

4: vancouver_trees

<!----------------------------------------------------------------------------->

1.2 One way to narrowing down your selection is to *explore* the datasets. Use your knowledge of dplyr to find out at least *3* attributes about each of these datasets (an attribute is something such as number of rows, variables, class type...). The goal here is to have an idea of *what the data looks like*.

*Hint:* This is one of those times when you should think about the cleanliness of your analysis. I added a single code chunk for you below, but do you want to use more than one? Would you like to write more comments outside of the code chunk?

<!-------------------------- Start your work below ---------------------------->

### *apt_buildings*
Acquired courtesy of The City of Toronto's Open Data Portal. It currently has 3455 rows and 37 columns.

```{r}
### EXPLORE HERE ###
head(apt_buildings)
```
### *cancer_sample*
Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.

```{r}
### EXPLORE HERE ###
head(cancer_sample)
```

### *building_permits*
Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 20680 rows and 14 columns.

```{r}
### EXPLORE HERE ###
head(building_permits)
```

### *steam_games*
Acquired courtesy of Kaggle. It currently has 40833 rows and 21 columns.

```{r}
### EXPLORE HERE ###
head(steam_games)
```

### *vancouver_trees*
Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 146611 rows and 20 columns.

```{r}
### EXPLORE HERE ###
head(vancouver_trees)
```


<!----------------------------------------------------------------------------->

1.3 Now that you've explored the 4 datasets that you were initially most interested in, let's narrow it down to 2. What lead you to choose these 2? Briefly explain your choices below, and feel free to include any code in your explanation.

<!-------------------------- Start your work below ---------------------------->
1: *building_permits*

2: *vancouver_trees*

The reason why I want to choose these 2 is, they have both categorical and numerical data. And their dimensions is relatively large, 
*building_permits* with dimensions of
```{r}
dim(building_permits)
```
*vancouver_trees* with dimensions of
```{r}
dim(vancouver_trees)
```

<!----------------------------------------------------------------------------->

1.4 Time for the final decision! Going back to the beginning, it's important to have an *end goal* in mind. For example, if I had chosen the `titanic` dataset for my project, I might've wanted to explore the relationship between survival and other variables. Try to think of 1 research question that you would want to answer with each dataset. Note them down below, and make your final choice based on what seems more interesting to you!

<!-------------------------- Start your work below ---------------------------->
My final decision is: 

**1: vancouver_trees**

**reasoning**: *vancouver_trees* dataset has very rich amounts and types of variables. In general, there are at least 4 types of variable, including geographical info, tree's categorical info, tree's numerical info (size), and time info. So we can combine any two of the variables to discuss the relationship between them. 

For example, one research question I can think of is:

1: The relationship between tree's planted time and tree's category. Is there a trend?

<!----------------------------------------------------------------------------->

# Important note

Read Tasks 2 and 3 *fully* before starting to complete either of them. Probably also a good point to grab a coffee to get ready for the fun part!

This project is semi-guided, but meant to be *independent*. For this reason, you will complete tasks 2 and 3 below (under the **START HERE** mark) as if you were writing your own exploratory data analysis report, and this guidance never existed! Feel free to add a brief introduction section to your project, format the document with markdown syntax as you deem appropriate, and structure the analysis as you deem appropriate. Remember, marks will be awarded for completion of the 4 tasks, but 10 points of the whole project are allocated to a reproducible and clean analysis. If you feel lost, you can find a sample data analysis [here](https://www.kaggle.com/headsortails/tidy-titarnic) to have a better idea. However, bear in mind that it is **just an example** and you will not be required to have that level of complexity in your project.

# Task 2: Exploring your dataset (15 points)

If we rewind and go back to the learning objectives, you'll see that by the end of this deliverable, you should have formulated *4* research questions about your data that you may want to answer during your project. However, it may be handy to do some more exploration on your dataset of choice before creating these questions - by looking at the data, you may get more ideas. **Before you start this task, read all instructions carefully until you reach START HERE under Task 3**.

2.1 Complete *4 out of the following 8 exercises* to dive deeper into your data. All datasets are different and therefore, not all of these tasks may make sense for your data - which is why you should only answer *4*. Use *dplyr* and *ggplot*.

1.  Plot the distribution of a numeric variable.
2.  Create a new variable based on other variables in your data (only if it makes sense)
3.  Investigate how many missing values there are per variable. Can you find a way to plot this?
4.  Explore the relationship between 2 variables in a plot.
5.  Filter observations in your data according to your own criteria. Think of what you'd like to explore - again, if this was the `titanic` dataset, I may want to narrow my search down to passengers born in a particular year...
6.  Use a boxplot to look at the frequency of different observations within a single variable. You can do this for more than one variable if you wish!
7.  Make a new tibble with a subset of your data, with variables and observations that you are interested in exploring.
8.  Use a density plot to explore any of your variables (that are suitable for this type of plot).

<!-------------------------- Start your work below ---------------------------->
Question 1.  Plot the distribution of a numeric variable.
The only numeric variable in this data set is **diameter** and **height_range_id**

Distribution of **diameter**
```{r}
answer2.1.1 <- vancouver_trees %>%
    ggplot(aes(diameter)) +
   geom_histogram(binwidth = 10)
print(answer2.1.1)
```

As we can see, the *diameter* ranges [0, 400], with most of its value are within [0, 50]

Distribution of **height_range_id**
```{r}
answer2.1.2 <- vancouver_trees %>%
    ggplot(aes(height_range_id)) +
    geom_histogram(binwidth = 1)
print(answer2.1.2)
```

As we can see, the *height_range_id* ranges[0, 10]. The height group with most counts is group 1 and group 2. 

Question 3.  Investigate how many missing values there are per variable. Can you find a way to plot this?
```{r}
answer2.3.1 <- vancouver_trees  %>%
    summarise_all(list(~is.na(.))) %>%
    pivot_longer(everything(),
               names_to = "variables", values_to="missing") %>%
    count(variables, missing) %>%
    ggplot(aes(y=variables,x=n,fill=missing))+
    geom_col()
print(answer2.3.1)
```

As we notice here, only *longitude*, *latitude*, *date_planted* and *cultivar_name* have missing values. Specifically, for *date_planted* and *cultivar_name*, its missing value's counts exceeds 50% of its total counts. 

Question 2.  Create a new variable based on other variables in your data (only if it makes sense)

Combining with some knowledge in forestry, I had an intuition that *genus_name*, *species_name* and *cultivar_name* could be concated together to form a variable to distinguish tree's category. However, through the misssing values analysis above, we know that *cultivar_name*'s missing value counts exceeds 50%. So, I decided to concate only *genus_name* and *species_name*.

```{r}
answer2.2.1 <- vancouver_trees %>%
    mutate(genus_species=paste(genus_name,species_name,sep="_")) %>%
    select(tree_id:species_name,genus_species, cultivar_name:latitude)
head(answer2.2.1)
```

As you can see, the new column *genus_species* is after *genus_name* and *species_name* and before *cultivar_name*

Question 7.  Make a new tibble with a subset of your data, with variables and observations that you are interested in exploring.

After exploring each column, I decided to drop *civic_number*,*std_street*, *longitude*,	*latitude*. because for the geo info, we have *neighbourhood_name*, *on_street* etc to determine its actual physical location.

```{r}
vancouver_trees_sub <- answer2.2.1 %>%
    select(tree_id,genus_name:date_planted)
head(vancouver_trees_sub)
```

Question 4.  Explore the relationship between 2 variables in a plot.

Question 6.  Use a boxplot to look at the frequency of different observations within a single variable. You can do this for more than one variable if you wish!

Combining question 4 and question 6 above, I created several boxplots as below to see relationship between variables.

**diameter** X **neighbourhood_name**

```{r}
answer2.6.1 <- ggplot(vancouver_trees_sub, aes(diameter, neighbourhood_name)) + 
    geom_boxplot(width = 0.8)
print(answer2.6.1)
```

**height_range_id** X **neighbourhood_name**

```{r}
answer2.6.2 <- ggplot(vancouver_trees_sub, aes(height_range_id, neighbourhood_name)) + 
    geom_boxplot(width = 0.8)
print(answer2.6.2)
```

**genus_name** X **diameter**

```{r, fig.width=10,fig.height=11}
answer2.6.3 <- ggplot(vancouver_trees_sub, aes(diameter, genus_name)) + 
    geom_boxplot(width = 0.8)
print(answer2.6.3)
```

**genus_name** X **height_range_id**

```{r, fig.width=10,fig.height=11}
answer2.6.4 <- ggplot(vancouver_trees_sub, aes(height_range_id, genus_name)) + 
    geom_boxplot(width = 0.8)
print(answer2.6.4)
```

As we can see above, **genus_name** as tree's category is more relevant to tree's size (diameter and height_range_id)

<!----------------------------------------------------------------------------->

# Task 3: Write your research questions (5 points)

So far, you have chosen a dataset and gotten familiar with it through exploring the data. Now it's time to figure out 4 research questions that you would like to answer with your data! Write the 4 questions and any additional comments at the end of this deliverable. These questions are not necessarily set in stone - TAs will review them and give you feedback; therefore, you may choose to pursue them as they are for the rest of the project, or make modifications!

<!--- *****START HERE***** --->
Based on the observation above, I defined 4 research questions as below:

1. Will the number of count of planted trees change by time?

2. What's the relationship between the year/month we planted the trees and its tree's height? Is there any trend by time?

3. What's the relationship between the neighbourhood and its tree's size (height and diameter)? Is there any pattern, such as some neighbourhood prefer planting bigger or taller trees?

4. What's the possible factor that contributes to the **curb** variable?

# Task 4: Process and summarize your data (13 points)

From Task 2, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions.

### 1.1 (10 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question.

Ensure that the output of each operation is printed!

**Summarizing:**

1.  Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2.  Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3.  Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4.  Based on two categorical variables, calculate two summary statistics of your choosing.

**Graphing:**

5.  Create a graph out of summarized variables that has at least two geom layers.
6.  Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.
7.  Make a graph where it makes sense to customize the alpha transparency.
8.  Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->
#### Research Question 1. Will the number of count of planted trees change by time?

First, in order to make the time variable *date_planted* more easier to be plotted and summarized, I separated it to *month* and *year* variable.

lubricate library is introduced here to make the date format transformation easier.
```{r}
library(lubridate, warn.conflicts = FALSE)
```

```{r}
vancouver_trees_time <- vancouver_trees_sub  %>%
                        mutate(month = floor_date(date_planted, unit="month")) %>%
                        mutate(year = format(as.Date(date_planted), "%Y"))%>%
                        drop_na(month)
head(select(vancouver_trees_time, c("date_planted", "year", "month")))
```

##### Summarizing: Compute the number of observations for at least one of your categorical variables.

```{r}
vancouver_trees_month_count <- vancouver_trees_time %>% 
  group_by(month) %>% 
  summarise(tree_count = n())
head(vancouver_trees_month_count, n = 15)
```

To get the time trend of tree planted count, we can compute its number of observations for time variable. And I noticed that it seems like there is a cyclic trend of the tree planted count (the number goes up and down) , which we might explore more when we plot the chart. 

##### Graphing: Create a graph out of summarized variables that has at least two geom layers.
```{r}
vancouver_trees_month_count_plot <- vancouver_trees_time %>% 
                                    filter(month > '2015-01-01') %>% 
                                    ggplot() +
                                #     geom_bar() + 
                                    geom_line(aes( x = month), stat="bin", binwidth=1) + 
                                    scale_x_date(NULL, date_labels = "%m-%Y", breaks = "month") +
                                    coord_flip()
print(vancouver_trees_month_count_plot)
```

The graph reflects the relationship between date and tree planted count. From the graph above, we noticed that there is a cyclic pattern of the tree planted count by month. And 1 cycle is roughly 1 year. Specifically, the month with larger tree planted count coincides with springtime (March to May). While during the winter month (Nov. to Feb.), people tends to plant less. This findings conform to our common sense. 


#### Research Question 2. What's the relationship between the year/month we planted the trees and its tree's height? Is there any trend by time?

##### Summarizing: Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.

*height_range_id* is an integer variable, with its definition of 0-10 for every 10 feet (e.g., 0 = 0-10 ft). So it's more like a catergorical value. To evaluate the relationship between time and tree's height, I calculated the **mean** of *height_range_id* across *year*.
```{r}
vancouver_trees_year_height <- vancouver_trees_time %>%
                              group_by(year) %>%
                              summarise(height_range_id_mean = mean(height_range_id, na.rm = TRUE))
head(vancouver_trees_year_height, n = 40)
```

we noticed that, the average of tree's height decreases year by year

##### Graphing: Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.

First, I transformed the data type of *year* and *height_range_id*, to make plotting easier
```{r}
vancouver_trees_time$height_range_id <- as.character(as.numeric(vancouver_trees_time$height_range_id))
vancouver_trees_time$year <- as.numeric(as.character(vancouver_trees_time$year))
```
```{r}
vancouver_trees_time_height_plot <- vancouver_trees_time %>% 
    ggplot(aes(x = year)) + 
    geom_line(aes(color = height_range_id),stat="bin",binwidth=1) 
print(vancouver_trees_time_height_plot)
```

The above graph describe the relationship between *year* and *planted_tree_count*, colored by *height_range_id*. Noticed that the *planted_tree_count* varies a lot, it's harder to distinguish the trend of smaller values. So I decided to apply log to *count*

```{r}
vancouver_trees_time_height_plot <- vancouver_trees_time %>% 
    ggplot(aes(x = year)) + 
    geom_line(aes(color = height_range_id),stat="bin",binwidth=1) +
    scale_y_continuous(trans='log2')
print(vancouver_trees_time_height_plot)
```

From the plot, we can determine that the tree's count of some smaller *height_range_id*, such as 0, 1 increases, and the larger ones, such as 2 - 6 decreases over the year. So, we can concludes that, people tend to plant shorter tree in the recent 10 years.

#### Research Question 3. What's the relationship between the neighbourhood and its tree's size (height and diameter)? Is there any pattern, such as some neighbourhood prefer planting bigger or taller trees?

##### Summarizing: Create a categorical variable with 3 or more groups from an existing numerical variable.

Since height has been transformed to categorical variable, we can also transform diameter, so that height and diameter are in the same range.

From Task 2, we know that *height_range_id* ranges [0, 10], while *diameter* ranges[0, 400], so I create a categorical variable *diameter_level*, which ranges [0, 14] as below
```{r}
vancouver_trees_diameter <- vancouver_trees_sub %>%
    mutate(diameter_level = cut(diameter,
                       breaks=c(0, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 100, 200, 300, 400),
                       labels=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14))) %>%
    mutate(diameter_level = as.numeric(as.character(diameter_level)))
head(vancouver_trees_diameter$diameter_level)
```

To evaluate the relationship between *neighbourhood_name* and tree's size, I calculated the **mean** of both *height_range_id* and *diameter_level* across **neighbourhood_name*.

```{r}
vancouver_trees_geo_size <- vancouver_trees_diameter %>%
  group_by(neighbourhood_name) %>%
  summarise(across(c("height_range_id", "diameter_level"), ~ mean(.x, na.rm = TRUE)))
print(vancouver_trees_geo_size)
```

##### Graphing: Create a graph out of summarized variables that has at least two geom layers.

To plot the *height_range_id* and *diameter_level* side by side for easier comparison, I used pivot_longer() to make the data "tidier".
```{r}
vancouver_trees_geo_size <- vancouver_trees_geo_size %>% 
                            pivot_longer(c(`height_range_id`, `diameter_level`), 
                             names_to = "size_type", 
                             values_to = "value")
print(vancouver_trees_geo_size)
```
```{r}
vancouver_trees_geo_size_plot <- vancouver_trees_geo_size %>% 
    ggplot(aes(x = neighbourhood_name, y = value , fill= size_type)) +
    geom_bar(stat="identity", position="dodge") + 
    coord_flip()
print(vancouver_trees_geo_size_plot)
```
In general, we noticed that *height_range_id* and *diameter_level* seems to have a linear relationship, when the trees is taller, the diameter is greater, which makes sense. And combining with *neighbourhood_name*. We can conclude that the neighbourhood with taller and bigger trees are 

1) West Point Grey

2) Shaughnessy

3) Kitsilano

4) Dunbar

5) Arbutus

They are mostly located in West Vancouver. 

#### Research Question 4. What's the possible factor that contributes to the **curb** variable?

##### Summarizing: Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.

To explor the relationship between *curb* and *street_side_name*, I calculated **percentage** of *curb* value ("Y" or "N") across *street_side_name*.

```{r}
vancouver_trees_stree_side_curb_percent <- vancouver_trees_sub %>% 
                                pivot_wider(names_from = curb,
                                values_from = curb,
                                values_fn = \(x)1,
                                values_fill = 0)%>%
                                group_by(street_side_name)%>%
                                summarise(across(N:Y, ~scales::percent(mean(.x), 0.01)))
head(vancouver_trees_stree_side_curb_percent)
```

We can notice that for Greenway, the trees 100% don't have curb, while for others type of street_side, the trees mostly have curb. Specifically, for Bike Road, the trees 100% have curb. It makes sense that for normal street, the trees should come with curb, while for Greenway the trees don't necessarily have ones.

##### Graphing: Create a graph out of summarized variables that has at least two geom layers.

To plot the *street_side_name* and *curb_percentage* in stack for easier comparison, I used pivot_longer() to make the data "tidier".
```{r}
vancouver_trees_stree_side_curb_percent <- vancouver_trees_stree_side_curb_percent %>% 
                            pivot_longer(c(`N`, `Y`), 
                             names_to = "curb", 
                             values_to = "value")
print(vancouver_trees_stree_side_curb_percent)
```
```{r}
vancouver_trees_stree_side_curb_percent_plot <- vancouver_trees_sub %>%
              count(street_side_name, curb) %>%       
              group_by(street_side_name) %>%
              mutate(percentage= prop.table(n) * 100) %>%
              ggplot() + aes(street_side_name, percentage, fill=curb) +
              geom_bar(stat="identity") +
              coord_flip()
print(vancouver_trees_stree_side_curb_percent_plot)
```

The graph perfectly illustrated our reasoning above in Summarize Section.

<!----------------------------------------------------------------------------->

### 1.2 (3 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!-------------------------- Start your work below ---------------------------->

Based on the exploration in Task 4 above, I refined my research proposal below:

#### Reseach Question 1. Will the number of count of planted trees change by time?

We noticed from the graph that there is a cyclic pattern of the tree planted count by month. And 1 cycle is roughly 1 year. Specifically, the month with larger tree planted count coincides with springtime (March to May). While during the winter month (Nov. to Feb.), people tends to plant less. This finding conform to our common sense. 

#### Reseach Question 2. What's the relationship between the year/month we planted the trees and its tree's height? Is there any trend by time?

We can determine from the summary and plot that, the tree's count of some smaller *height_range_id*, such as 0, 1 increases, and the larger ones, such as 2 - 6 decreases over the year. So, we can concludes that, people tend to plant shorter tree in the recent 10 years.

#### Reseach Question 3. What's the relationship between the neighbourhood and its tree's size (height and diameter)? Is there any pattern, such as some neighbourhood prefer planting bigger or taller trees?

In general, we noticed that *height_range_id* and *diameter_level* seems to have a linear relationship, when the trees is taller, the diameter is greater, which makes sense. And combining with *neighbourhood_name*. We can conclude that the neighbourhood with taller and bigger trees are 

1) West Point Grey

2) Shaughnessy

3) Kitsilano

4) Dunbar

5) Arbutus

They are mostly located in West Vancouver. 

#### Reseach Question 4. What's the possible factor that contributes to the **curb** variable?

We can learn from the graph that, for Greenway, the trees 100% don't have curb, while for others type of street_side, the trees mostly have curb. Specifically, for Bike Road, the trees 100% have curb. It makes sense that for normal street, the trees should come with curb, while for Greenway the trees don't necessarily have ones.

<!----------------------------------------------------------------------------->

### Attribution

Thanks to Icíar Fernández Boyano for mostly putting this together, and Vincenzo Coia for launching.